---
title: "p8105_hw3_cn2604"
output: github_document
date: "2022-10-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r load_library, echo=FALSE, message=FALSE}
library(tidyverse)
library(p8105.datasets)
library(patchwork)
library(ggridges)
```

# Problem 2
```{r clean_file_1, echo=FALSE, message=FALSE}
acce_df = read.csv("./data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "activity_number",
    names_prefix = "activity_",
    values_to = "activity_counts"
  ) %>% 
  mutate(
    week_type = factor(case_when(
      day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday") ~ "weekday",
      day %in% c("Saturday", "Sunday") ~ "weekend",
      TRUE ~ "unknown"
      )
    ),
    .after = day
  )

day_df = acce_df %>% 
  group_by(week, day_id) %>% 
  summarize(
    day_activity = as.integer(sum(activity_counts))
  )

```
The dataframe after cleaning processes has `r ncol(acce_df)` variables and `r nrow(acce_df)` observations. It includes the week and day variables with an indication of either weekday or weekend. Each one minute interval is turned into a variable named activity_number and activity_counts is the new variable for original activity data.

```{r show_table, echo=FALSE, message=FALSE}
knitr::kable(day_df)
day_df %>% 
  ggplot(aes(x=day_id, y=day_activity, color = week)) + geom_point() + geom_line()
```


# Problem 3
```{r load_file_2, echo=FALSE, message=FALSE}
data("ny_noaa")
```
The raw dataframe of ny_noaa has `r ncol(ny_noaa)` variables and `r nrow(ny_noaa)` observations. It has weather station id, date of observation, precipitation, snowfall, snow depth, and maximum and minimum temperatures. For data of precipitation, snowfall, and snow depth, there are `r sum(is.na(ny_noaa$prcp))`, `r sum(is.na(ny_noaa$snow))`, `r sum(is.na(ny_noaa$snwd))` missing values. For tmax and tmin, `r sum(is.na(ny_noaa$tmax))` and `r sum(is.na(ny_noaa$tmin))` of the observations are missing, which loss about a half of helpful information for further analysis. 
```{r clean_file_2, echo=FALSE, message=FALSE}
noaa_df = ny_noaa %>% 
  separate(date, into = c("year", "month", "day"), sep = '-') %>% 
  rename(c("prcp_mm" = "prcp", "snow_mm" = "snow", "snwd_mm" = "snwd"))

clean_snow = na.omit(noaa_df$snow_mm)
find_mode = function(v) {
 uniqv = unique(v)
 uniqv[which.max(tabulate(match(v, uniqv)))]
}

```
1. While most of the observations for snowfall are missing, `r find_mode(clean_snow)` mm is the most commonly seen value in the remaining observation. It is reasonable because snowfall is an occasional event in that area and no snow should be observed in normal days.
```{r df_for_plot, echo=FALSE, message=FALSE}
jan_tmax = noaa_df %>% 
  filter( month == "01") %>% 
  select(id, year, tmax) %>% 
  na.omit() %>% 
  mutate(
    tmax = as.numeric(tmax),
    year = as.numeric(year)
    ) %>%
  group_by(id, year) %>% 
  summarize(
    mean_tmax_jan = mean(tmax)
  )

july_tmax = noaa_df %>% 
  filter( month == "07") %>% 
  select(id, year, tmax) %>% 
  na.omit() %>% 
  mutate(
    tmax = as.numeric(tmax),
    year = as.numeric(year)
    ) %>% 
  group_by(id, year) %>% 
  summarize(
    mean_tmax_july = mean(tmax)
  )

```
```{r plots, echo=FALSE, message=FALSE}
jan_tmax %>% 
  ggplot(aes(x=year, y=mean_tmax_jan)) + geom_point()
```


